---
title: "Categorical data homework 1"
author: "mercy"
date: "November 16, 2020"
output: html_document
---

##  Question 1a
Find and sketch the likelihood function. Give the ML estimate of pi

```{r}

n=20

L = function(pi_,x){
  term = (pi_^x)*((1-pi_)^(1-x))
  return(prod(term))
}

x=rep(1,n)
pi_ = seq(0,1,by=0.001)
plot(pi_, sapply(pi_,L,x=x),type='l')
abline(v = 1, col = "gray60")
```

## Question 1b 

wald test and 95% CI
wald test statistics =
$$ z=  \hat \theta - \theta_0 / SE_0(\hat\theta)$$

wald confidence interval =

$$
\hat \theta \pm z_{\alpha/2} SE(\hat \theta)
$$
The SE can be estimate by 

$$
\mathit{SE} = \sqrt \frac{\hat \theta(1-\hat \theta)}{n}
$$

Given that $\hat \theta = 1$  and $n=20$
our SE will be equal to 

$$
\mathit{SE} = \sqrt \frac{1(1-1)}{20}
$$
This is not sensible and hence our wald test statistics as well as the confidence intervals willnot be sensible.

## Question 1c
score test and the corresponding p-value. Construct a 95% score confidence interval.
The score Test statistic is given equating the second derivative of the likelihood to zero.
Therefore we have that, 
$$
z = \frac{\hat \theta - \theta_0}{\sqrt { \theta_0(1- \theta_0)/n}}
$$

Therefore the the score test statistic equals
$$
z = \frac{1 - 0.5}{\sqrt { 0.5(1- 0.5)/20}} = 4.4721
$$

The p-value is given as
```{r}
2*pnorm(-abs(4.4721))
```
The ver small value of the p-value indicate that there is a statistically significance difference between the $\theta$ values from 0.5

The CI is given by
$$
\hat \pi \Bigg(\frac{n}{n + z^2_{\alpha/2}} \Bigg)  
+ \frac{1}{2} \Bigg( \frac{ z^2_{\alpha/2}} {n + z^2_{\alpha/2} }           \Bigg) \\
\pm z^2_{\alpha/} \sqrt {\frac{1}{n + z^2_{\alpha/2}} \bigg[  \hat \pi (1 - \hat\pi) \bigg(\frac{n}{n + z^2_{\alpha/2}}\bigg)+ \bigg(  \frac{1}{2}\bigg) \bigg(  \frac{1}{2}\bigg)\bigg( \frac{z^2_{\alpha/2}}{n+z^2_{\alpha/2}}  \bigg)        
\bigg]}


$$
This results to....to be calculated or check below

```{r}

 prop.test
 
 
```



# 1d likelihood ratio test and its CI

The likelihood ratio test statistic is given by 

$$
-2(L_0 - L_1) = 2\bigg( xlog\frac{\hat\pi}{\pi_0} + (n-x)log\frac{1-\hat\pi}{1-\pi_0} \bigg)

$$
12.0412
The 95% likelihood ratio confidence interval is the set of $\pi_0 $ for which the likelihood ratio statistic
$$
-2(L_0 - L_1) \le X^2_1(0.05)
$$
```{r}
1-pchisq (12.04, df = 1)
```

# question E
Construct an exact binomial test and 95% confidence interval.
check pclab1 for the test \\
For the  95% CI

# question 3 
\begin(verbatim)
attach(data)
chisq <- chisq.test(data)
chisq
\end(verbatim)




$$
X-squared = 25.086, df = 4, p-value = 4.835e-05
$$


attach(data)
chisq <- chisq.test(data)
chisq
the row and the column variables are statistically significantly associated (p-value < 0.05)

\begin(verbatim)

chisq$observed
round(chisq$expected,2)
#Pearson residuals
round(chisq$residuals, 3)

\end(verbatim)

Because the chi sq statistics can only tell if there is an association between the two variables, we go further to do a pearson residual to check the strengh of the association.

# Question 4a
No, 11.4 is actually the ratio of odds and not the probabilities. In addition if these two this would be case if independence holds
The right interpretation would be that the odds of survival for females was 11.4 times the odds for males.
#4b 
OR for survival , $\theta = 11.4$
Odds for frmales who survived = 2.9

Proportion for female adults who survived will be
$$
\frac {p_1}{1-p_1} = 2.9 \\
p_1 = 2.9 (1-p_1)\\
p_1 + 2.9p_1 = 2.9\\
p_1 = 2.9/3.9 = 0.7436

$$
Proportion of males who survided will be 
$$
\theta = \frac {odds of females who survived}{odds of males who survived}\\
11.4 = 2.9/ males who survived\\
odds males who survived = 2.9/11.4 = 0.2544
$$
let p_2 be the propoetion of males who survived, 
$$
\frac{p_2}{1-p_2} = 0.2544\\
p_2 = 0.2028


$$











 

